\chapter{Introduction}

As the \ac{IoT} continuously grows in importance~\cite{reviewIoT}, collecting high-frequency data from smart meters offers unprecedented opportunities for operational efficiency~\cite{SmartMetersApplications}. A potential use case for high-frequency data is \ac{DR} systems. \ac{DR} systems utilize smart meter readings to actively balance the supply and demand of electricity on the grid. For instance, utility providers can offer dynamic pricing or incentive payments to encourage consumers to reduce electricity use during periods of high wholesale market prices~\cite{DemandResponseSurvey}. Additionally, by analyzing the resulting consumption shifts, energy providers can adjust generation schedules to balance energy production levels, ultimately improving overall grid efficiency~\cite{DemandResponseBenefits}.

However, collecting detailed consumption data from smart meters poses significant privacy risks. High-resolution smart meter readings can reveal sensitive household characteristics regarding marital status, employment status, income, or the number of residents~\cite{RevealingHouseholdCharacteristicsSM}. Moreover, daily behavioral patterns, including residents' presence, sleep schedules, or meal times, can be inferred from the recorded energy profiles. When combined with metadata about specific appliances, the exposed behavioral routines may be exploited by adversaries for malicious purposes~\cite{InferringPersonalInformationFromDRS,AshgarSurveyPaper}.

Consequently, protecting consumer privacy is not merely a technical necessity for user safety, but a strict legal requirement. Within the European Union, the General Data Protection Regulation (GDPR), effective since 2018, establishes strict rules for processing, storing, and managing personal data of individuals. The primary goal of the GDPR is to give individuals more control over their personal information and ensure that utility companies manage digital records responsibly~\cite{GDPR}.

A naive approach might suggest removing explicit private information (e.g., names or addresses) from published datasets to avoid re-identification attacks. Yet, it is well-established that simply removing direct identifiers remains insufficient for protecting user privacy~\cite{Sweeney_SimpleDemographics}. Foundational research into infamous real-world examples, such as the de-anonymization of the Netflix Prize dataset, has proven that individuals can be easily re-identified by linking seemingly innocuous data points known as \acp{QI}~\cite{NetflixPaper}.

To counter identity disclosure, formal privacy models like $k$-anonymity were developed for static datasets~\cite{Sweeney_kanon}. However, traditional $k$-anonymity models are often too computationally intensive for the real-time, zero-delay requirements of modern smart grid streams. The need for low-latency processing has led to the development of stream-specific algorithms, such as $z$-anonymity, which are designed to provide lightweight, continuous privacy protection~\cite{zanonDataStreams}.

While $z$-anonymity provides core privacy guarantees, the effectiveness of the algorithm and the resulting data utility heavily depend on data preprocessing techniques. Because modern \ac{IoT} networks are distributed and multi-layered~\cite{reviewIoT}, consisting of smart meters, gateways, and a Central Entity, as described precisely in Section~\ref{sec:SMandIoT}, relying on a conventional centralized privacy model requires placing a high degree of trust in the Central Entity collecting the raw, sensitive data~\cite{dezent2025}. To reduce central trust requirements, the data stream can be preprocessed closer to the source prior to the formal privacy check. This thesis investigates how shifting specific tasks, such as generalizing, aggregating, and prefiltering, to the Edge and Fog layers affects the resulting privacy-utility-performance trade-off. As the findings of this research demonstrate, performing data generalization at the Edge layer can significantly improve data availability compared to raw data baselines while reducing excessive data suppression. Furthermore, temporal aggregation effectively lowers communication overhead, whereas local prefiltering at the Fog layer offers notable bandwidth savings, albeit with some impact on global data availability.

To systematically evaluate the impact of deploying the selected preprocessing techniques across different architectural layers, the present work establishes a baseline scenario where raw data is processed exclusively by the Snapshot $z$-anonymity algorithm (as described in Section~\ref{sec:zanonsnapshot}) at the Central Entity. The baseline is then compared against variations where data transformations are initially applied at the Edge (on smart meters) or at the Fog (gateways) Layer. By measuring the effects on data availability, information loss, and network performance, the research provides a clear framework for understanding the trade-offs involved in moving computation closer to the end user. All experiments are grounded in a realistic context using the \textit{SmartMeter Energy Consumption Data in London Households}~\cite{londonSmartMeter2014} dataset. 

This thesis is structured as follows. Chapter~2 introduces the necessary background. Chapter~3 reviews the related literature. Chapter~4 presents the System Model. Chapter~5 describes the methodology, while Chapter~6 presents the experimental results. Chapter~7 discusses the results, and Chapter~8 concludes the thesis.