\pagestyle{empty} % Removes page numbering

\begin{center}
    \vspace*{1cm}
    {\Large Generative AI Usage} 
    \vspace{1cm}
\end{center}

During the preparation of this thesis, generative AI tools (Google Gemini) were used in a supportive capacity. These tools assisted with literature exploration, programming tasks, and language refinement.

For literature-related work, AI was used to obtain general topic overviews and clarify specific technical concepts. While the tools provided initial summaries, all cited sources were independently identified, reviewed, and verified in their original form by the author. No citations were included based solely on AI suggestions.

For programming, AI provided support in refining Python-based simulation framework. The tools assisted in ensuring better code modularity, debugging syntax errors, and assisting with the implementation of data visualizations. All underlying logic and experimental scenarios were designed by the author, and all final code was manually reviewed and validated through successful execution.

For language improvement, AI was used to enhance clarity and grammatical correctness. The tools provided support by suggesting alternative phrasing for draft sentences to ensure an appropriate academic tone. In all cases, the underlying content, interpretations, and conclusions originated from the author. All AI-assisted text was critically reviewed and manually edited to ensure accuracy.


The author takes full responsibility for the final content of this thesis, and all AI-assisted outputs were critically evaluated.

\newpage